---
title: "vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(QuiPTsim)
```

# Introduction

QuiPTsim is a framework generating matrices of n-gram occurences of given sequences, which will be further used to assess feature selection algorithm QuiPT, which allows very fast feature filtering.

# Sequence generation

generate_single_motif 
 : function generates motif from an alphabet
 : `n` - number of elements sampled from an alphabet
 : `d` - total sum of gaps
 : `motifProbs` - (optional) probabilites of alphabet elements
 
```{r}
generate_single_motif(1:4)
generate_single_motif(c("a", "b", "c"))
generate_single_motif(1:4, n = 6, d = 2, motifProbs = c(0.7, 0.1, 0.1, 0.1))
```

In our simulations, we have decided to use $n=4$ and $d=6$, which gives maximum motif length being 10.


---

generate_motifs 
 : function returns list of motifs (results of `generate_single_motif()`)
 : function asserts if such motifs can be injected to a sequence (sequence length = 10)

```{r}
print(generate_motifs(alphabet = letters[1:6], n_motif = 3))
```

---

simulate_single_sequence
 : generates sequence of elements from alphabet with replacement
 : (optional) alphabet probabilites can be specified
 
```{r} 
print(simulate_single_sequence(len = 10,
                         alphabet = letters[1:6]))

# we can also generate sequence based on sequence elements probabilites
probs <- runif(6)
probs <- probs / sum(probs)
print(probs)
print(simulate_single_sequence(len = 10,
                         alphabet = letters[1:6],
                         seqProbs = probs))

```
---

add_motifs 
 : function injects motifs to a sequence

each sequence is returned with attributes: 
- list of motifs
- list of according masks
 
Important thing about motif injection is the fact, that multiple motifs can overlap each other. 

```{r}
alph <- letters[1:6]

example_sequence <- simulate_single_sequence(len = 10, alph)
motifs_list <- generate_motifs(alph, 2)
print(example_sequence)
print(motifs_list)
injected_sequence <- add_motifs(sequence = example_sequence,
                                motifs = motifs_list)
print(injected_sequence)
```

---

simulate_sequences 
 : function generates sequences (both positive & negative)
 : matrix of sequences is returned
 : motifs are injected to a fraction of sequences
 : following attributes are added:
 * list of motifs
 * list of masks
 * target variable (flag if motifs have been injected to a sequence)
 
```{r}
numberOfSequences <- 10
sequencesLength <- 10
alphabet <- letters[1:4]
motifs <- generate_motifs(alph, 3)

# Although motifs list contain 3 independent motifs
# only one will be injected
# Since fraction = 0.2, motifs will be injected only into 2 sequences
simulate_sequences(numberOfSequences,
                   sequencesLength, 
                   alphabet,
                   motifs,
                   n_motifs = 1,
                   fraction = 0.2)

```

---

generate_sequences 
 : wrapper for a `simulate_sequences()` and place where n-grams are counted
 : n-grams are counted by `count_multimers()` function from `seqR` package, which allows us to simulate data at terrific speed
 
Utilization of `seqR::count_multimers()` is tricky, as we are supposed to deliver vector of all k-mers' lengths along with vector of gaps for each corresponding k-mer length.

---

# Simulation data

create_simulation_data
 : function is a wrapper of functions implemented in sequence generation block.
 
```{r}
# alphabet elements
alph <- letters[1:6]

# number of replication for each set of parameters (sequence length, number of motifs, etc.)
reps <- 10

# Number of sequences to be generated
n_seq <- c(10)

# Sequences' length
l_seq <- c(10)

# Number of motifs that will be injected to a sequence
n_motifs <- c(1, 2)

# results will be saved in this location
path <- "./"

# triggers if matrices will be saved (debugging purposes)
save_files <- FALSE

# title of simulation
title <- "SEQ"

results <- create_simulation_data(reps, n_seq, l_seq, n_motifs, alph, path, title, save_files)
results
```

On top of above example, `create_simulation_data` allows us to specify probability vectors for both motifs and sequences. Furthermore, MD5 sums are being collected right after object save. 

---

# Cosine similarity

euclidean_norm
 : returns euclidean norm of a vector

```{r}
x <- runif(5)
print(x)
print(euclidean_norm(x))
```

---
 
cosine_similarity
 : return cosine similarity of two vectors

```{r}
x <- runif(5)
y <- runif(5)
print(x)
print(y)
print(cosine_similarity(x, y))
```
---

generate_probs
 : return two vectors of probabilites with given cosine similarity

```{r}
size <- 6
cs <- 0.8

probabilites <- generate_probs(size, cs)
print(probabilites)

print(cosine_similarity(probabilites[1, ], probabilites[2, ]))
```

